---
title: "R-problemset_econometrics"
output:
  html_document: default
  pdf_document: default
date: "2022-10-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This is the R-problem set 1 for Econometrics (Fall 2022), New York University
Author: Carlos Figueroa (cdf5579)

Let's begin.

```{r}
#lets load some libraries first

library(dplyr)
library(magrittr)
library(readxl)
library(tinytex)
library(data.table)
library(tibble)
library(ggplot2)



```


lets display the data first

```{r}

data = data.frame(read.csv(file = "bertrand2004.csv", header = TRUE , sep =","))

head(data)
```

QUESTION 1:--------------------------------------------------------------------

A. What is the interpretation of C¯ in this situation? Hint: this is a 0/1 variable like a coin
flip, so what is the mean of a Bernoulli?

Since C (which stands for a person recieving a call back or not) is between zero and one, it will behave similarly to a bernoulli distribution. That would mean that C bar (the sample mean of C) is going to be equal to p, or the probability of receiving a call back. We can get this p for different subgroups and the group in general.

B. What about M¯ ? Hint: the same as the previous hint.

Since M also follows a bernoulli distribution, its sample mean is going to be p, so such mean is going to be equal to the probability of being a man in this dataset.


c. Let C¯M and C¯F represent the call back rates for males and females respectively. Set up a hypothesis test for whether or not there is discrimination on sex.

Given that we stated that C bar represents the sample distribution, and the probability of recieving a callback, we will want to analyze C bar (M) and C bar (F) to see if one group has higher chances of recieving a callback than the other.

Then, our hypothesis testing will be for a two tail rest:

H0: C¯M = C¯F then C¯M - C¯F = 0
H1: C¯M != C¯F 

If we disprove H0, we will say that there is discrimination on sex in this particular dataset.


Here, in this section, we are going to see if both groups have similar qualifications in order to make this hypothesis test viable. If men or women are more qualified, that will change our view of the results of our hypothesis test.
 
QUESTION 2---------------------------------------------------------------------

Calculate the mean years of experience and education by sex

```{r}

exp_male = mean(data[data$sex == 'm', 'yearsexp'])
exp_female = mean(data[data$sex == 'f', 'yearsexp'])


educ_male = mean(data[data$sex == 'm', 'education'])
educ_female = mean(data[data$sex == 'f', 'education'])


print("Mean of years of Experience for males")
exp_male
print("Mean of years of Experience for females")
exp_female

print("Mean of years of Education for males")
educ_male
print("Mean of years of Education for females")
educ_female
```

Female have more experience than male, and male have more education. But only by little, they are pretty close. Let's test if they are significally different.


QUESTION 3---------------------------------------------------------------------


Calculate the standard errors and do a t-test (by hand or using the test commands) for
whether years of experience or education are different by sex

```{r}

print("Standard Deviations for Male (years of experience)")
sd(data[data$sex == 'm', 'yearsexp'])
print("Standard Deviations for Female (years of experience)")
sd(data[data$sex == 'f', 'yearsexp'])
print("Standard Deviations for Male (years of education)")
sd(data[data$sex == 'm', 'education'])
print("Standard Deviations for Female (years of education)")
sd(data[data$sex == 'f', 'education'])

#lets confirm our results with proper t-test package

t.test(data[data$sex == 'm', 'yearsexp'], data[data$sex == 'f', 'yearsexp'], var.equal = TRUE)


t.test(data[data$sex == 'm', 'education'], data[data$sex == 'f', 'education'], var.equal = TRUE)

```


QUESTION 4---------------------------------------------------------------------

Now, using the same commands (or any other commands you prefer), calculate the difference
in call back rates by sex.


```{r}

print("Probability of recieving a callback by being a male")
mean(data[data$sex == 'm', 'call'])


print("Probability of recieving a callback by being a female")
mean(data[data$sex == 'f', 'call'])

```

QUESTION 5---------------------------------------------------------------------

Finally, do a t-test to see if these differences are statistically significant. You may do this by
hand, with the t.test command, or with any other command.


```{r}

t.test(data[data$sex == 'm', 'call'], data[data$sex == 'f', 'call'], var.equal = TRUE)

```
Not that different indeed.

PART 2 - - - - - - - - - -- -- - - -- - - -- - - -- - - - - - - - -- - - - - - -


Regression analysis


QUESTION 6---------------------------------------------------------------------

First, load the dataset just as before. For this exercise, you will also need the ggplot2 library
loaded.


```{r}
df = data.frame(read.csv(file = "stocks.csv", header = TRUE , sep =","))


head(df)
```

QUESTION 7---------------------------------------------------------------------

Second, let’s plot the data. Using ggplot2 make a scatter plot of the returns of Tesla as
the dependent variable and the returns of market as the independent variable. To this, load
the ggplot2 package. For the command, here is a rough schema (that you have to fill in
yourself):

The first part tells R to make a ggplot object, using which dataset, x, and y variables are
specified. The second part says to create a scatter plot.



```{r}

ggplot(data = df, aes(x = ret_sp500,y = ret_tsla)) + geom_point()

```

QUESTION 8---------------------------------------------------------------------

 Now let’s add a best fit line. To do this we copy and paste the same opening as above with
an additional command telling R to add the line:



```{r}


ggplot(data = df, aes(x = ret_sp500,y = ret_tsla)) + geom_point() + geom_smooth(method = 'lm',se=FALSE)


```


QUESTION 9---------------------------------------------------------------------

Now let’s run a regression to calculate the beta of Tesla stock. To do this we need to use the
lm command. The schematic as below, where you need to fill in the y, x, and data entries
yourself:


```{r}

lm(ret_tsla ~ ret_sp500, data = df)

```

QUESTION 10---------------------------------------------------------------------


What is the β of Tesla stock? Calculate the β for each of the other three stocks in the data.
Fill in a table that looks like the one at the end of the homework.

1.334472

```{r}

lm(ret_tsla ~ ret_sp500, data = df)

lm(ret_aapl ~ ret_sp500, data = df)

lm(ret_msft ~ ret_sp500, data = df)

lm(ret_f ~ ret_sp500, data = df)



```
So the table will go:

For Tesla: 1.334472, Apple: 1.1987471, Microsoft: 1.1883818, Ford: 1.0576747  


QUESTION 11---------------------------------------------------------------------



The results from the base lm command are a bit bare. To learn more, we use the summary
command. To do this, we need to save our model and then feed it to the summary command:


Once you’ve run this command, you can report the variance of the error term from your
regressions, σ2U, by squaring the residual standard error. Run the command for each of the
four stocks in our data—add these results to your table containing the β’s. Which stock has
the most volatile returns net of market comovement (highest σ2U)?.


```{r}


model = lm(ret_tsla ~ ret_sp500, data = df)

summary(model)

model_1 = lm(ret_aapl ~ ret_sp500, data = df)

summary(model_1)

model_2 = lm(ret_msft ~ ret_sp500, data = df)

summary(model_2)

model_3 = lm(ret_f ~ ret_sp500, data = df)

summary(model_3)

```
```{r}

0.03354^2

0.01233^2

0.009469^2

0.01822^2

```

For Tesla: 1.334472 and 0.001124932, Apple: 1.1987471 and 0.0001520289, Microsoft: 1.1883818 and 8.966196e-05, Ford: 1.0576747 and 0.0003319684

The highest u variance is for tesla, with 0.001124932. So Tesla is the most volatile stock.

